---
params:
  author: 
    label: "Report author:"
    value: "Cliff Clavin"
  email: 
    label: "Author's email address:"
    value: "cliff_clavin@fws.gov"
  phone:
    label: "Author's phone #:"
    value: "555-867-5309"
  o_dir: 
    label: "Full file path to occurrence data:"
    value: "C:/Users/mcobb/Desktop/r_webinar/docs/mccrea/demo/data/raw"
    input: text
  s_dir: 
    label: "Full file path to sample site data:"
    value: "C:/Users/mcobb/Desktop/r_webinar/docs/mccrea/demo/data/raw"
    input: text
  hab_data: 
    label: "Input habitat covariate raster (.asc):"
    value: 
    input: file
author: |
  | `r params$author`
  | Cheers National Wildlife Refuge
  | National Wildlife Refuge System, Boston Region
  | U.S. Fish and Wildlife Service
  | `r params$email`
  | `r params$phone`
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: true
    includes:
      in_header: header.tex
      before_body: before_body.tex
---


```{r setup, include=FALSE}

# Load packages
library(raster)
library(unmarked)
library(knitr)
library(rmarkdown)

knitr::opts_chunk$set(echo = FALSE, results = "asis", warning = TRUE)

# Get user parameters
s_dir <- params$s_dir
o_dir <- params$o_dir
hab_data <- params$hab_data

```


```{r data1, echo=FALSE, results="hide", cache=FALSE, message=FALSE, warning=FALSE, error=FALSE}

o.files <- list.files(o_dir, pattern = "occurrence", full.names = TRUE)
# o.files <- list.files("./data/raw", pattern = "occurrence", full.names = TRUE)
o.years <- as.numeric(sapply(o.files, extract.year))
# o.files <- o.files[o.years <= year]
o.years <- as.numeric(sapply(o.files, extract.year))

s.files <- list.files(s_dir, pattern = "sample_site", full.names = TRUE)
# s.files <- list.files("./data/raw", pattern = "sample_site", full.names = TRUE)
s.years <- as.numeric(sapply(s.files, extract.year))
# s.files <- s.files[s.years <= year]
s.years <- as.numeric(sapply(s.files, extract.year))

##miss.o <- !(s.years %in% o.years)
##miss.s <- !(o.years %in% s.years)
years <- intersect(o.years, s.years)
year <- max(years)
occu.files <- o.files[o.years %in% years]
site.files <- s.files[s.years %in% years]

## Load survey year data
nyears <- length(years)
nlogical <- ifelse(nyears==1, FALSE, TRUE)

# Load spatial data and plot survey map
habcov <- raster(hab_data)
# habcov <- raster("../resources/data/geodata/habcov.asc")
nsites <- length(habcov)

# Load site and occurrence data
sitedfs <- lapply(site.files, read.csv)
occudfs <- lapply(occu.files, read.csv)

## Process and format data
sdmpData <- process.data(occudfs, sitedfs, years, habcov)
class(sdmpData) <- c("sdmpData", "list")

# save(sdmpData,file = paste0("../data/final/species_occurrence_data_", year, ".RData"))
figcap <- paste0("Plot of site-specific naive occurrence summed over ", nyears,
                 " survey years and represented as scaled blue circles. ",
                 "Cell colors represent habitat covariate values.")
```

\newpage

## Project description

The Species Disribution Monitoring Project is an annual refuge
Inventory and Monitoring survey began in 2011 to monitor species
occurrence and the relationship between occurrence and the
distribution of habitat across a National Wildlife Refuge.
The survey consists of conducting multiple surveys for a species presence.
The survey design as described in the Refuge Species
Distribution Monitoring Project Protocol requires the surveys be
conducted at a random sample of 20 sites each year. With a given year,
each site is surveyed for species presence every day for a consecutive
5-day period, each day (i.e., occasion) constituting an independent sample of occurrence data. Sites
are equal-area polygons (i.e., grid cells) as defined by a lattice
overlaid on the Refuge boundary (Fig. \@ref(fig:surveymap)) which
divides the refuge into `r nsites` unique sites. Observers are
instructed to search each cell to which he/she is assigned with the
same intensity to standardize detection by survey effort.  Detection
is known to be affected by visibility conditions and, therefore, a
categorical visibility covariate (i.e., good vs poor) is collected for
each survey occasion in each year to be used in data analysis. Habitat
quality is a known factor affecting the probability of species
occurrence. Because habitat quality across the the refuge does not
substantially vary over time (i.e., year to year), a static spatial
data set measuring habitat quality at the level of the survey site
(Fig. \@ref(fig:habitatmap)) is used in the analysis to account for variation in occurrence due to habitat quality.

\newpage

## Overall project summary

The Species Occurrence Project began in 2011 and surveys
have been conducted each year up to this year,
`r max(years)`. This section provides summary information for the
entire duration of the project.

\newpage

### Survey area

The survey area corresponds to the entire refuge and is divided into `r nsites`
equal-area polygons, each constituting a single survey site with a
unique site ID (Fig. \@ref(fig:surveymap)). Each site is characterized by a
measure of habitat quality measured on the the real number
scale (Fig. \@ref(fig:habitatmap)).

```{r surveymap, echo=FALSE, results="hide", fig.cap="Plot of the survey sites (i.e., grid cells) with site IDs and refuge boundary (blue line)."}

#print(plot.surveymap(habcov))
verts = as.vector(sp::bbox(habcov))
rasterpts = data.frame(raster::rasterToPoints(habcov))
raster::plot(raster::rasterToPolygons(habcov), axes = TRUE, xlab = "X", ylab= " Y", asp = 1)
text(rasterpts$x, rasterpts$y, label = 1:nrow(rasterpts), cex = 0.75)
polygon(x = verts, y = verts[c(1,3,4,2)], lwd = 2, border = "blue")
```

```{r habitatmap, echo=FALSE, results="hide", fig.cap="Plot of the habitat covariate values (scaled colors) at survey sites with site IDs listed at cell centroids."}

#print(plot.habitatmap(habcov))
rasterpts = data.frame(rasterToPoints(habcov))
raster::plot(habcov, xlab = "X", ylab = "Y", asp = 1)
text(rasterpts$x, rasterpts$y, label = 1:nrow(rasterpts), cex = 0.75)
```

```{r habcovhist, echo=FALSE, results="hide", fig.cap="Histogram of habitat covariate measured at all survey sites within study area."}

#print(plot.habcov.hist(habcov))
hist(values(habcov), breaks = seq(-3,3,0.2), ylim = c(0, 20),
     main = "", xlab = "Habitat covariate")
```

\newpage

### Survey data

```{r condition1, echo=FALSE, results="hide", cache=FALSE, message=FALSE, warning=FALSE, error=FALSE}

naiveoccus <- sapply(sdmpData$NaiveOccu, mean)
if(nyears==1){
    text1 <- paste0("To date, 1 covariate has been collected for detection and 1 covariate has been collected",
                    " for species occurrence. The naive proportion of sites occupied in ",
                   years, " was ", naiveoccus, ".")
} else {
    text1 <- paste0("To date, 1 covariate has been collected for detection and 1 covariate has been collected",
                    " for species occurrence. The naive proportions of sites occupied each year ranged from ",
                    min(naiveoccus), " to ", max(naiveoccus),
                    "(Fig. \\@ref(fig:naiveoccuplot)).")
}
```

```{r print1, results="asis",echo=FALSE}
cat(text1)
```

```{r naiveoccuplot, eval=nlogical, echo=FALSE, results="asis", fig.cap="Plot of naive proportion of sites occupied each year."}

if(nyears>1){
    plot(years, naiveoccus, ylim = c(0,1), type = "b", pch = 16, frame.plot = FALSE,
         axes = FALSE, ylab = "Naive proportion of sites occupied", xlab = "Year")
    axis(2)
    axis(1, at = years)
}
```


\newpage

## Individual year summaries

This section contains year-specific summary information ...

\newpage

```{r echo=FALSE, results="asis", message=FALSE, warning=FALSE, error=FALSE}

# Directory path for figures
#figdir.path <- "../products/reports/annual_qc/figure/"
dir.create("figure")
figdir.path <- "./figure/"

dummy <- mapply(annual.summaries, 
                occu = sdmpData$occuDataList, 
                site = sdmpData$siteDataList,
                year = sdmpData$years, 
                unmkd = sdmpData$unmarkedDataList, 
                naiveoccu = sdmpData$NaiveOccu,
                MoreArgs = list(habcov = sdmpData$habcovData, figdir.path = figdir.path),
                SIMPLIFY = FALSE)
```

\clearpage

## Appendix

\clearpage

### R session info

```{r echo=FALSE, results="asis", message=FALSE, warning=FALSE, error=FALSE}
sessionInfo()
```

---
title: "Species Distribution Monitoring Project: Annual data summary and quality
    control report for `r max(years)`"
---
